---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "`r Sys.getenv('USER')`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: true
---

```{r setup}

library(scales) # for comma_format
library(tidyverse)
library(dplyr)
library(tidyr)
library(janitor)
library(cowplot)

```

## Summarize results of bulk processing for outlier lead accuracy project

This data comes from the CKCC1 outlier results.

## Load data

```{r}

gathered_results_raw <- read_tsv("2020_12_03/outlier_probability_output_CKCC_2020_12_04.tsv.gz") 

gathered_results <- gathered_results_raw %>%
  select(Sample, Gene, Type, Expression, 
         Threshold, MEND_depth = `Mend depth`,
         effective_length = `Gene length`,
         Mean_Accuracy = `Probability of being an outlier`) %>%
  mutate(`Percent difference` = 100 * ((abs(Expression - Threshold)) / Expression)) %>%
  drop_na(Mean_Accuracy)

# Results including NAs

gathered_results_with_nas <- gathered_results_raw %>%
  select(Sample, Gene, Type, Expression, 
         Threshold, MEND_depth = `Mend depth`,
         effective_length = `Gene length`,
         Mean_Accuracy = `Probability of being an outlier`) %>%
  mutate(`Percent difference` = 100 * ((abs(Expression - Threshold)) / Expression))

```

## Add bin info

```{r}

expression_bin_boundaries <- c(0, 1, 3, 5, 7, 10, 20)
mend_depth_bin_boundaries <- c(0, seq(2, 46, 4), Inf)*1E6
length_bin_boundaries <- c(0, 100, 450, 800, 1500, 4000, Inf)


make_bin_names <- function(boundaries = 1:10){
  raw_names <- paste0(boundaries, "-", lead(boundaries)) %>% str_replace("-Inf", "+")
  raw_names[1:(length(boundaries)-1)]
}
  

expression_bin_names <- make_bin_names(expression_bin_boundaries)
mend_depth_bin_names <- make_bin_names(mend_depth_bin_boundaries/1e6)
length_bin_names <- make_bin_names(length_bin_boundaries)

bulk_results <- gathered_results %>%
  mutate(Expression_bin = cut(Expression, expression_bin_boundaries, 
                              labels = expression_bin_names,
                        include.lowest = TRUE),
         MEND_bin = cut(MEND_depth, mend_depth_bin_boundaries, labels = mend_depth_bin_names),
         length_bin = cut(effective_length, length_bin_boundaries, labels = length_bin_names,
                                      include.lowest = TRUE)
         )

# Results including NAs

bulk_results_with_nas <- gathered_results_with_nas %>%
  mutate(Expression_bin = cut(Expression, expression_bin_boundaries, 
                              labels = expression_bin_names,
                        include.lowest = TRUE),
         MEND_bin = cut(MEND_depth, mend_depth_bin_boundaries, labels = mend_depth_bin_names),
         length_bin = cut(effective_length, length_bin_boundaries, labels = length_bin_names,
                                      include.lowest = TRUE)
         )

```

## Summary stats on bins

```{r}

print("How many detected outlier genes total are being analyzed?")
nrow(bulk_results)

print("How many bins are possible?")
(length(expression_bin_boundaries)-1) * (length(mend_depth_bin_boundaries)-1) * (length(length_bin_boundaries)-1)

bin_check <- bulk_results %>%
  group_by(Expression_bin, MEND_bin, length_bin) %>%
  summarize(n_entries_per_bin = n())

bin_check_with_nas <- bulk_results_with_nas %>%
  group_by(Expression_bin, MEND_bin, length_bin) %>%
  summarize(n_entries_per_bin = n())

print("How many bins contain entries in this data?")
nrow(bin_check_with_nas)

print("How many bins have between 1 and 99 results?")
length(which(bin_check_with_nas$n_entries_per_bin > 0 & bin_check_with_nas$n_entries_per_bin < 100))

```

## Confirm the absence of NA values

```{r}

sum(is.na(bulk_results$Mean_Accuracy))

```

## Look at bin sizes

```{r}

# Make sure (depth, expression, length) bins are relatively similar sizes

all_bins <- bulk_results %>% tabyl(MEND_bin, Expression_bin, length_bin)

all_bins

```

## How many outliers had >95% accuracy?

```{r}

print("How many outliers had >95% accuracy?")
length(which(bulk_results$Mean_Accuracy > 0.95))

print("What percent of outliers had >95% accuracy?")
100*length(which(bulk_results$Mean_Accuracy > 0.95))/nrow(bulk_results)

```

## Spectrum of accuracies by depth, expression, and length

```{r}

plot2 <- ggplot(bulk_results, aes(x=Mean_Accuracy)) +
  geom_histogram() +
  facet_wrap(~MEND_bin) +
  ggtitle("Spectrum of accuracies by MEND depth")

plot2

plot3 <- ggplot(bulk_results, aes(x=Mean_Accuracy)) +
  geom_histogram() +
  facet_wrap(~Expression_bin) +
  ggtitle("Spectrum of accuracies by expression level")

plot3

plot4 <- ggplot(bulk_results, aes(x=Mean_Accuracy)) +
  geom_histogram() +
  facet_wrap(~length_bin) +
  ggtitle("Spectrum of accuracies by effective gene length")

plot4

```

## StatBin2 function def

```{r}

# StatBin2 allows depiction of empty bins as blank instead of a horizontal line:
# https://stackoverflow.com/questions/57128090/remove-baseline-color-for-geom-histogram
StatBin2 <- ggproto(
  "StatBin2", 
  StatBin,
  compute_group = function (data, scales, binwidth = NULL, bins = NULL, 
                            center = NULL, boundary = NULL, 
                            closed = c("right", "left"), pad = FALSE, 
                            breaks = NULL, origin = NULL, right = NULL, 
                            drop = NULL, width = NULL) {
    if (!is.null(breaks)) {
      if (!scales$x$is_discrete()) {
        breaks <- scales$x$transform(breaks)
      }
      bins <- ggplot2:::bin_breaks(breaks, closed)
    }
    else if (!is.null(binwidth)) {
      if (is.function(binwidth)) {
        binwidth <- binwidth(data$x)
      }
      bins <- ggplot2:::bin_breaks_width(scales$x$dimension(), binwidth, 
                                         center = center, boundary = boundary, 
                                         closed = closed)
    }
    else {
      bins <- ggplot2:::bin_breaks_bins(scales$x$dimension(), bins, 
                                        center = center, boundary = boundary, 
                                        closed = closed)
    }
    res <- ggplot2:::bin_vector(data$x, bins, weight = data$weight, pad = pad)

    # drop 0-count bins completely before returning the dataframe
    res <- res[res$count > 0, ] 

    res
  })

```

# Figure 2

```{r figure-2}

# to get hex codes from palette
# library(RColorBrewer)
# brewer.pal(12, "Set1")

type_colors = c(Up = "#E41A1C", Down = "#377EB8")

split_by_quartile <- function(bulk_results_in_fn, col_name_fn = Expression, this_suffix = " Log2(TPM+1)"){
  #  bulk_results_in_fn <- bulk_results_for_ckcc_accuracy_figure; col_name_text = "Depth"; this_suffix = "M MEND reads"
  col_name_text <- rlang::ensym(col_name_fn)
  # Upper quartile
 bulk_results_bit <- lapply(c("Up", "Down"), function(this_type){
    these_bulk_results <- bulk_results_in_fn %>% filter(Type == this_type)
    this_threshold <- these_bulk_results %>% summarize(this_q = quantile(!!col_name_text, 0.75)) %>% pull(this_q) %>% unique
    bulk_results_upper_bit <- these_bulk_results %>% 
      filter(!!col_name_text >= this_threshold) %>%
      mutate(set = paste0(rlang::as_label(col_name_text),  "_UQ"),
             set_type_label = paste0(">=", round(this_threshold, 1), this_suffix, ", n=", label_comma()(n())))
  # Lower quartile
    this_threshold <- these_bulk_results %>% summarize(this_q = quantile(!!col_name_text, 0.25)) %>% pull(this_q) %>% unique
    bulk_results_lower_bit <- these_bulk_results %>% 
      filter(!!col_name_text <= this_threshold) %>%
      mutate(set = paste0(rlang::as_label(col_name_text), "_LQ"),
             set_type_label = paste0("<=", round(this_threshold, 1), this_suffix, ", n=", label_comma()(n())))
   these_bits <- bind_rows(bulk_results_upper_bit, bulk_results_lower_bit) 
  }) %>% bind_rows()
  return(bulk_results_bit)
}

bulk_results_for_ckcc_accuracy_figure <- 
  bulk_results %>% 
  mutate(Depth = MEND_depth/1e6,
         Length = effective_length/1e3,
         Difference = `Percent difference`)

ckcc_accuracy_figure_data_v2 <- bind_rows(
split_by_quartile(bulk_results_for_ckcc_accuracy_figure, col_name = Expression, this_suffix = " Log2(TPM+1)") %>% mutate(fac1 = "Expression"),
split_by_quartile(bulk_results_for_ckcc_accuracy_figure, col_name = Depth, this_suffix = "M MEND reads") %>% mutate(fac1 = "Depth"),
split_by_quartile(bulk_results_for_ckcc_accuracy_figure, col_name = Length, this_suffix = "kb") %>% mutate(fac1 = "Length"),
split_by_quartile(bulk_results_for_ckcc_accuracy_figure, col_name = Difference, this_suffix = "%") %>% mutate(fac1 = "Difference")
)

tabyl(bulk_results_for_ckcc_accuracy_figure$Type)

```


```{r figure-2}

# to get hex codes from palette
# library(RColorBrewer)
# brewer.pal(12, "Set1")

type_colors = c(Up = "#E41A1C", Down = "#377EB8")

acc_overview <- ggplot(bulk_results, aes(x=Mean_Accuracy, color=Type)) +
  geom_histogram(fill="white", stat = StatBin2) +
  facet_wrap(~Type, scales="free") +
  theme_bw() +
  scale_color_manual(values = type_colors) +
  theme(legend.position="none") +
  xlab("") +
  ylab("Gene\nexpression\nmeasurements")

```

## Generate figures

```{r figure-2g}

boxplot_n_labels <- ckcc_accuracy_figure_data_v2 %>%
   select(set, set_type_label, Type, set_type_label, fac1) %>%
  distinct()


all_boxplots <- ggplot(ckcc_accuracy_figure_data_v2, aes(y=set, x=Mean_Accuracy, fill = Type)) +
  geom_boxplot(outlier.size = 0.5, width = 0.5) +
  geom_text(data = boxplot_n_labels, 
            aes(x=0.05, y=set, label = set_type_label), hjust = 0, vjust = -1.5, size = 3) +
  theme_bw() +
  scale_fill_manual(values = type_colors) +
  scale_color_manual(values = type_colors) +
  facet_grid(fac1~Type, scales = "free_y") +
#    facet_wrap(~Type_label, ncol = 2) +
  theme(legend.position="none") +
  ylab("") +
  xlab("Probability")

all_boxplots

```


```{r figure-2f, fig.height=7, fig.width=7}

combined_image <- plot_grid(acc_overview, all_boxplots, 
                            ncol = 1, 
                            rel_heights = c(1,2), 
                            align = "v",
                            axis = "lr",
                            labels = "AUTO")
combined_image

ggsave("../figures/ckcc_accuracy_figure.png", combined_image, height=7, width=7)

```

## Where does the small peak in the left tail for up outliers in fig 2A come from?

```{r}

# Outliers with probability < 0.05
low_acc_ups <- subset(bulk_results, Mean_Accuracy < 0.05)

# How many are there?
nrow(low_acc_ups)

# What percent are short & low exp?
100*length(which(low_acc_ups$Expression_bin == "0-1" & low_acc_ups$length_bin == "0-100"))/nrow(low_acc_ups)

```

# Why are accuracies lower for upper percent difference quartile than lower percent difference quartile?

```{r}

pd2 <- figure1_data_v2 %>%
  filter(set == "Difference<2.2%")

pd100 <- figure1_data_v2 %>%
  filter(set == "Difference>100%")

print("pd2 up summary of expression & accuracy")
summary(pd2$Mean_Accuracy[pd2$Type == "Up"])
summary(pd2$Expression[pd2$Type == "Up"])

print("pd2 down summary of expression & accuracy")
summary(pd2$Mean_Accuracy[pd2$Type == "Down"])
summary(pd2$Expression[pd2$Type == "Down"])

print("pd100 up summary of expression & accuracy")
summary(pd100$Mean_Accuracy[pd100$Type == "Up"])
summary(pd100$Expression[pd100$Type == "Up"])

print("pd100 down summary of expression & accuracy")
summary(pd100$Mean_Accuracy[pd100$Type == "Down"])
summary(pd100$Expression[pd100$Type == "Down"])

```

## Why are the accuracies lower for longer down outliers?

```{r}

long_downs <- figure1_data_v2 %>%
  filter(Type == "Down" & set == "Length>5.7kb")

short_downs <- figure1_data_v2 %>%
  filter(Type == "Down" & set == "Length<0kb")

print("long down summarys")
summary(long_downs$Expression)
summary(long_downs$`Percent difference`)
summary(long_downs$MEND_depth)

print("short down summarys")
summary(short_downs$Expression)
summary(short_downs$`Percent difference`)
summary(short_downs$MEND_depth)

```
## Why does fig 2A look like sunshine & rainbows, while fig 2B looks like tornadoes & hailstorms?

```{r}

ggplot(figure1_data_v2, aes(y=Type, x=Mean_Accuracy, fill = Type)) +
  geom_boxplot(outlier.size = 0.5) +
  theme_bw() +
  scale_fill_manual(values = type_colors) +
  scale_color_manual(values = type_colors) +
  facet_wrap(~Type, ncol = 2) +
  theme(legend.position="none") +
  ylab("")

ggplot(figure1_data_v2, aes(y=Type, x=Mean_Accuracy, fill = Type)) +
  geom_violin(outlier.size = 0.5) +
  theme_bw() +
  scale_fill_manual(values = type_colors) +
  scale_color_manual(values = type_colors) +
  facet_wrap(~Type, ncol = 2) +
  theme(legend.position="none") +
  ylab("")

ggplot(subset(figure1_data_v2, set == "Expression<0.1 Log2(TPM+1)" & Type == "Up")) +
  geom_histogram(aes(Mean_Accuracy)) +
  theme_bw()

```



